---
title: "MLB Win Rate Prediction"
author: "sandeul"
date: "2018/4/12"
output1: html_document
---


```{r setup, include=FALSE}
library(data.table)
library(corrplot)
library(dplyr)
library(car)
library(gvlma)
library(ggplot2)
library(usdm)
library(PerformanceAnalytics)
library(lm.beta)
library(caret)

pitchingDF <- fread(input = "data/pitching.txt",
                    sep = ",",
                    header = TRUE)

battingDF <- fread(input = "data/batting.txt",
                   sep = ",",
                   header = TRUE)
str(pitchingDF)

pitchingDF <- pitchingDF[,-c(1,4,5,6,7,12,36)]

TeamDF <- merge(battingDF, pitchingDF, by = c("Year", "Tm"))
TeamDF <- TeamDF[Year < 2018 ,]

```


#**EDA(Exploratory data analysis)**
### 1.Variables
```{r}
colnames(TeamDF)
```

### 2. Sample Size
```{r}
nrow(TeamDF)
```

### 3. Dependent Variable(y)
```{r}
summary(TeamDF$winrate)
boxplot(TeamDF$winrate)
```


### 4. Independent Variable(x)
```{r}
str(TeamDF)
colSums(is.na(TeamDF))
```


### 5. Train / Test
```{r}
# train 70% / test 30%
set.seed(9943)
inTrain <- createDataPartition(y=TeamDF$winrate, p=0.7, list=FALSE)
trainDF <- TeamDF[inTrain , ]
testDF  <- TeamDF[-inTrain , ]
```


# Correlation Analysis
```{r}
analysis.pit <- trainDF[,c("winrate", "ERA", "SHO", "SV", "IP", "Hpit", "Rpit", "ER", "HRpit", "BBpit", "IBBpit", "SOpit", "HBPpit", "BK", "WP", "BF", "WHIP", "H9", "HR9", "BB9", "SO9", "SO/W")]

analysis.bat <- trainDF[,c("winrate", "#Bat", "R/G", "G", "PA", "AB", "Rbat", "Hbat", "2Bbat", "3Bbat", "HRbat", "RBI", "SB", "CS", "BBbat", "SObat", "BAbat", "OBP", "SLG", "OPS", "TB", "GDP", "HBPbat", "SH", "SF", "IBBbat", "LOB")]


 chart.Correlation(analysis.pit, histogram=TRUE, pch=4)
 chart.Correlation(analysis.bat, histogram=TRUE, pch=4)

```


#### %Emergency Situation%
```{r}
TeamDF[TeamDF$G < 160 , Year]
divideG <- function(args) {
  return(args/TeamDF$G)
}

values <- c("PA", "AB", "Rbat", "Hbat", "2Bbat", "3Bbat", "HRbat", "RBI", "SB", "CS", "BBbat", "SObat", "HBPbat", "TB", "SH", "SF", "IBBpit", "CG", "SHO", "SV", "IP", "Hpit", "Rpit", "ER", "HRpit", "BBpit", "IBBpit", "SOpit", "HBPpit", "BK", "WP", "BF", "GDP")
for (i in values) {
  TeamDF[,i] <- divideG(TeamDF[,i, with = FALSE])
}

```


```{r include=FALSE}
trainDF <- TeamDF[inTrain , ]
testDF  <- TeamDF[-inTrain , ]
```


```{r}
analysis.pit <- trainDF[,c("winrate", "ERA", "SHO", "SV", "IP", "Hpit", "Rpit", "ER", "HRpit", "BBpit", "IBBpit", "SOpit", "HBPpit", "BK", "WP", "BF", "WHIP", "H9", "HR9", "BB9", "SO9", "SO/W")]

analysis.bat <- trainDF[,c("winrate", "#Bat", "R/G", "G", "PA", "AB", "Rbat", "Hbat", "2Bbat", "3Bbat", "HRbat", "RBI", "SB", "CS", "BBbat", "SObat", "BAbat", "OBP", "SLG", "OPS", "TB", "GDP", "HBPbat", "SH", "SF", "IBBbat", "LOB")]

 chart.Correlation(analysis.pit, histogram=TRUE, pch=4)
 chart.Correlation(analysis.bat, histogram=TRUE, pch=4)

```


```{r}
# Batter Data
corrplot.mixed(cor(analysis.pit), p.mat = abs(cor(analysis.pit[,-1])), sig.level = 0.5, tl.pos = "lt", tl.col = "black")
# R/G, Rbat, RBI, SLG, OPS, TB, Hbat, PA, BAbat, BBbat, HRbat, 2Bbat, AB, SF, CS, SObat, SH > 0.5


# Pitcher Data
corrplot.mixed(cor(analysis.bat), p.mat = abs(cor(analysis.bat[,-1])), sig.level = 0.5, tl.pos = "lt", tl.col = "black")
# ERA, Rpit, ER, Hpit, HRpit, BBpit, SOpit, SO9, BF, H9, SHO, HBPpit, HR9, BB9, SO/W, IP >0.5


# Variable Selection
analysis.bat <- trainDF[,c("winrate", "#Bat",  "3Bbat", "SB",  "OBP", "GDP", "HBPbat", "IBBbat", "LOB")]
analysis.pit <- trainDF[,c("winrate", "SV", "IBBpit", "BK", "WP", "WHIP")]

```

# **Linear regression**

```{r}
# Collective Data
analysis2 <- cbind(analysis.bat, analysis.pit[,-1])
# Multiple Regression Analysis
lm.result <- lm(winrate~ .,data =  analysis2)
# Analysis Result
summary(lm.result)

```

### 1. Selecting Independent Variable
```{r}

# Removing meaningless data ( WP, 3Bbat, HBPbat, BK,  IBBpit)
analysis2 <- analysis2[,c("winrate", "SV", "WHIP", "#Bat", "SB",  "OBP", "GDP", "IBBbat", "LOB")]

lm.result <- lm(winrate~ .,data =  analysis2)
summary(lm.result)


# Stepwise variable selection
step(lm.result, direction = "both")

lm.result <-  lm(formula = winrate ~ SV + WHIP + `#Bat` + SB + OBP + GDP + IBBbat + LOB, data = analysis2)

# No change

```

### 2. Multicollinearity

```{r}
car::vif(lm.result)
```

### 3. Residual Analysis
```{r}
gvlma.result <- gvlma(lm.result)
summary(gvlma.result)
par(mfrow = c(2,2))
plot(lm.result)
```

### 4. Final Model
```{r}
beta.result <-lm.beta(lm.result)
beta.result$coefficients %>% abs() %>% sort() %>% round(digits = 6)

# LOB, IBBbat, `#Bat`, SB, GDP 제거
lm.result <-  lm(formula = winrate ~ SV + WHIP  + OBP, data = analysis2)
summary(lm.result)

beta.result <-lm.beta(lm.result)
beta.result$coefficients %>% abs() %>% sort() %>% round(digits = 6)

chart.Correlation(lm.result$model, histogram=TRUE, pch=4)

```

### 5. Final Model Residual Test
```{r}
gvlma.result <- gvlma(lm.result)
summary(gvlma.result)
par(mfrow = c(2,2))
plot(lm.result)
```

# **Predictive Model**

```{r}
# Final Model
finalModel <- lm(formula = winrate ~ SV + WHIP + OBP, data = analysis2)

# prediction test
pred <- predict(finalModel, 
                newdata = testDF,
                interval = "predict")
pred <- as.data.table(pred)

# predicted result
pred.final <- pred
test.final <- testDF[, winrate]

pred.final$winrate <- test.final

pred.final$result <- ifelse((pred.final$lwr <= pred.final$winrate) & (pred.final$winrate <= pred.final$upr), 1, 0 )
pred.final
table(pred.final$result)
```

# **Result** 

## winrate =  0.20574 + 0.25852&lowast;SV - 0.50251&lowast;WHIP + 2.80361&lowast;OBP
</br>
```{r}
```
## Standarized regression coefficient
```{r}
lm.beta::lm.beta(lm.result)
```


* SV   : Save  
* WHIP : Walks hits per inning pitched  
* OBP  : On base percentage
```{r}
```
## Predictive Rate : (419 / 439) &lowast; 100 = 95.4%   
## Model Prediction resulted in a high rate 95.4%.  
</br></br>